---
title: "modelbuilder Package Tutorial"
author: "<a href=http://handelgroup.uga.edu>Andreas Handel</a>"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
    css: vignettestyle.css
vignette: >
  %\VignetteIndexEntry{DSAIDE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: vignetterefs.bib
---

**Most features are not yet implemented, this is a roadmap!**

## Summary
Future home of the modelbuilder vignette


## Package description

The package consists of several functions that allow a user to create and analyze compartmental dynamical models in a graphical user interface setting without having to write code.


## Intended audience and goal of the package

The audience for this app are individuals interested in dynamical models who are new to coding, but nevertheless want to develop and analyze such computer models. 


## Installing the package 
The package can be installed from CRAN or Github. See the documentation on the [package page](https://ahgroup.github.io/modelbuilder/) for more details.

Package installation is a one-time process, unless R itself is being upgraded/reinstalled. Note that the package depends on other packages, which will also be installed as needed.


## Starting the package

Every time a new R/Rstudio session is started, the package needs to be loaded: 
```{r, eval=FALSE, echo=TRUE}
library('modelbuilder')
```
followed by starting the main menu for the package:
```{r, eval=FALSE, echo=TRUE}
modelbuilder()
```

This will bring up a graphical user interface from which everything is done. Once your are done using `modelbuilder`, simply close the graphical user interface.

## Building a model

The graphical user interface has two main tabs, one where you build the model and one where you analyze the model. You always start with the build tab. Here you can graphically build a model by adding variables, flows and parameters. As you build the model you will see a graphical representation of the model you are building. Once you're done building the model, and you're happy with it, you should click save model. This will save your model in an Rdata file in the location you specify.
Similarly, by pressing the save diagram button, you can save the diagram for the model as a PNG file in the location you want.

If you previously built a model, you can load it with the load model button. Once the model is loaded you can further modify it and save a modified version.

So far the model has not been implemented as a specific set of equations. You need to do that before you can simulate and analyze your model. You can choose what kind of implementation you want for the model, for instance a differential equation model or a stochastic model. After you have chosen your implementation, you can press the create model button. You will be asked where you want the code to be saved. This code is a fully working R function that you could use, run and analyze by yourself without using the modelbuilder package any further. You can also switch over to the analyze tab to graphically analyze this model.



## Analyzing a model
Once you have created a model, saved it, and produced a function for a specific implementation for instance an ODE function, you can use the analyze tab to simulate your model for different parameter values and graphically analyze the results.

By clicking the "export scenario" button, you can export the code that runs the model with the settings you chose and produces the figures shown in the GUI.


## Analyzing a model

The analyze tab is meant to give you some basic analysis tools. If you want to do more sophisticated analyses, for instance running a loop over a specific parameter, you will need to move on from the GUI and start using the underlying code produced by the modelbuilder package, and combine it with your own extra lines of code. The following shows an example of how this can be done.



## Further ideas

Have an import button that allows users to import models that are written in for instance SMBL

## References
