---
title: "modelbuilder Tutorial"
author: "<a href=https://www.andreashandel.com>Andreas Handel</a>"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
    css: vignettestyle.css
vignette: >
  %\VignetteIndexEntry{modelbuilder Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: vignetterefs.bib
---


## Summary
The modelbuilder package allows the building and analysis of compartmental dynamical systems models without the need to write code.


## Package description

The package provides a graphical user interface that allows a user to create and analyze compartmental dynamical models without having to write code. 


## Intended audience and goal of the package

The audience for this app are individuals interested in dynamical models who are new to coding, but nevertheless want to develop and analyze such computer models. 


## Installing the package 

The package has not yet been submitted to CRAN, therefore it currently needs to be installed from Github. To do so, you need the `remotes` package. If you don't have it, install it first. The following commands will get you up and running:

```{r, eval=FALSE, echo=TRUE}
install.packages('remotes')
remotes::install_github('ahgroup/modelbuilder')
```

Package installation is a one-time process, unless R itself is being upgraded/reinstalled. Note that the package depends on other packages, which will also be installed as needed.



## Starting the package

Every time a new R/Rstudio session is started, the package needs to be loaded: 
```{r, eval=FALSE, echo=TRUE}
library('modelbuilder')
```
followed by starting the main menu for the package:
```{r, eval=FALSE, echo=TRUE}
modelbuilder()
```

This will bring up a graphical user interface from which everything is done. Once your are done using `modelbuilder`, simply close the graphical user interface.


## Main menu

* The 'Main' tab allows you to load a previously created and saved model, which needs to be built by the modelbuilder package and stored in an .Rdata file. During loading, the package will check to ensure the model is valid. Alternatively, you can load a few example models. 
* To remove a currently loaded model, clear it. 
* If a model is loaded, you can export the R code that runs the model as either ODE, stochastic or discrete-time model. Note that this is a 'one-way' process, you can (currently) not edit the R code you receive and load it back into `modelbuilder`. Thus, this functionality is meant if you are ready to take matters in your own hand and modify the code beyond what can be done with the `modelbuilder` package.

## Building a model

* The 'build' tab allows you to either build a new model if currently no model is loaded, or modify an existing, currently loaded model. 
* You should probably first sketch out your model as a diagram (and/or equations) so you know which compartments, parameters and processes/flows to include.
* Use the add/remove buttons to add variables/flows/parameters to the model. 
* A few notes on model building: 
    - You need to fill all empty fields.
    - All variables need to start with an uppercase letter, all parameters need to start with a lowercase letter. Only letters and numbers are allowed. 
    - Flows can include variables, parameters and the following mathematical symbols: +-*/^() Other symbols (e.g. `sin` or `cos`) are not yet implemented.
    - If a flow indicates moving from one compartment to another, you need to write it twice. E.g. if susceptible move to infected, you need a `-bSI` outflow for the susceptibles and the corresponding `bSI` inflow for infected.
    - Every parameter used in one of the flows must be specified in the _model parameter_ section.

* Once you think your model is correct, press the _Make Model_ button. `modelbuilder` will check to make sure all the information you entered seems valid and follows the above listed rules. **That does not mean your model is correct/right!** It only means that the error checks couldn't find anything that made it look like an improperly defined `modelbuilder` model object. If there are errors, you will be told to fix them. Once no errors are found, you have the option to save your model as an .Rdata file. This is recommended if you want to later load and further edit or analyze it. If you don't save and `modelbuilder` closes for whatever reasons, your model is gone.
* Once your model has been created, you will see the equations displayed. Check them to make sure they look as you expect, of not modify your model. 
* NOT YET IMPLEMENTED:
  - A diagram for the model will be drawn
  - You can download the diagram and the equations as figures


## Analyzing a model

* To analyze a model, you first need to load one or create one with the build functionality. 
* nce you have loaded or created a model, you can use the analyze tab to simulate your model for different parameter values and graphically analyze the results. You can choose if you want to run the model as ODE, stochastic or discrete time model.
* NOT YET IMPLEMENTED:
  - Export code that runs the model with chosen settingsand produces the figure shown in the GUI.



## Roadmap

In addition to the not-yet implemented features stated above, the following are planned:

- Have import/export function to interface with SBML

Let us know if there are others you want to see implemented!


## References
