---
title: "modelbuilder Package Tutorial"
author: "<a href=http://handelgroup.uga.edu>Andreas Handel</a>"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
    css: vignettestyle.css
vignette: >
  %\VignetteIndexEntry{DSAIDE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: vignetterefs.bib
---


## Summary
The modelbuilder package allows the building and analysis of compartmental dynamical systems models without the need to write code.


## Package description

The package provides a graphical user interface that allows a user to create and analyze compartmental dynamical models without having to write code. 


## Intended audience and goal of the package

The audience for this app are individuals interested in dynamical models who are new to coding, but nevertheless want to develop and analyze such computer models. 


## Installing the package 
The package is currently only on Github. Install from there.

Package installation is a one-time process, unless R itself is being upgraded/reinstalled. Note that the package depends on other packages, which will also be installed as needed.



## Starting the package

Every time a new R/Rstudio session is started, the package needs to be loaded: 
```{r, eval=FALSE, echo=TRUE}
library('modelbuilder')
```
followed by starting the main menu for the package:
```{r, eval=FALSE, echo=TRUE}
modelbuilder()
```

This will bring up a graphical user interface from which everything is done. Once your are done using `modelbuilder`, simply close the graphical user interface.


## Building a model

The 'build' tab allows you to either build a new model, or modify an existing, loaded model (modify functionality not yet fully implemented). 

Under the build tab, you can graphically build a model by adding variables, flows and parameters. Once you finished specifying your model, you can generate it. This checks for errors, if none are present it will create the underlying model object and show the equations as set of differential equations for the underlying model.

Once you're done building the model, and you're happy with it, you should save model. This will save your model in an Rdata file in the location you specify.




## Analyzing a model
to analyze a model, you first need to load one (in the form of an Rdata file) that you or someone else previously built. 

Once you have created a model, saved it, and produced a function for a specific implementation for instance an ODE function, you can use the analyze tab to simulate your model for different parameter values and graphically analyze the results.
By clicking the "export scenario" button, you can export the code that runs the model with the settings you chose and produces the figures shown in the GUI.

Similarly, by pressing the save diagram button, you can save the diagram for the model as a PNG file in the location you want.

If you previously built a model, you can load it with the load model button. Once the model is loaded you can further modify it and save a modified version.



## Exporting model code
If you want to perform analyses beyond what you can do through the graphical interface, you have the option of exporting code for the model as R function. You can choose if you want code for an ODE, discrete-time, or stochastic model. 



## Further ideas

**Most features are not yet implemented, this is a roadmap!**


a model diagram (not yet implemented)

Have an import button that allows users to import models that are written in for instance SMBL

## References
